Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference

### 2024-05-17 paper/paper2_neo/paper.tex
Start of Gully edit of the paper prose
had to direct copy-paste text from VScode into the web to circumvent annoying git permissions issues

\section{Introduction}
Stellar spectra are exceedingly rich sources of information about the stars
which create them. They contain information about the fundamental properties
of the star such as its temperature and surface gravity, as well as
countless absorption lines reflecting the star's chemical composition. On 
top of that, extrinsic factors such as the radial velocity of the star or 
the rotation of the star itself further transform spectra. When these spectra
are observed, even the location of the observing instrument and the 
capabilities of the instrument itself introduce further modifications. 
In short, stellar spectra represent extremely complex, high-dimensional 
data that has gone through multiple transformations before reaching our 
eyes as a deceptively simple graph of flux versus wavelength.

Scientists want to extract as much information as possible from
a star's spectra to get back the properties that lie at the root of it all. 
However, this has proven to be no easy task. Spectroscopic surveys such as 
APOGEE develop their own in-house pipelines to extract stellar parameters
from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
effective \citep{APOGEE, ASPCAP}. However, they all work on the core assumption that the data to be 
analyzed is a list of pixels. In addition, the pipelines that surveys 
develop are typically closed-source and usually limited to the scope of the 
survey itself, causing a sort of fragmentation. This motivates the 
development of a more universal, open-source tool that can do what spectral 
analysis pipelines do, but in an alternative way.

>>>>

%token commit by gully!

\section{Introduction}
Stellar spectra are exceedingly rich sources of information about the stars
that produce them. Spectra encode fundamental properties such as temperature, surface gravity, and chemical composition via their
seemingly countless absorption lines. Extrinsic properties such as the stellar radial velocity or 
the projected equatorial rotation further transform spectra. Our observations of the spectra
degrade the otherwise pure astronomical signals, and so the resolution the observing instrument 
and other instrumental properties limit our ability to extract fundamental properties perfectly. 
Stellar spectra represent extremely complex, high-dimensional data that has gone through multiple transformations before reaching our detectors.
\authorcomment1{I like that you mention high-dimensional.  It recommend expanding on the meaning of high dimensional.  ``High'' in which sense?}

The challenge of extracting as much information as possible from
a stellar spectrum to get back the properties that lie at the root of it all. 
However, this has proven to be no easy task. Spectroscopic surveys such as 
APOGEE develop their own in-house pipelines to extract stellar parameters
from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
effective \citep{APOGEE, ASPCAP}. However, they all work on the core assumption that the data to be 
analyzed is a list of pixels. In addition, the pipelines that surveys 
develop are typically closed-source and usually limited to the scope of the 
survey itself, causing a sort of fragmentation. This motivates the 
development of a more universal, open-source tool that can do what spectral 
analysis pipelines do, but in an alternative way.
